---
title: "Vignette illustrating the usage of gscreend on simulated data"
author: "Katharina Imkeller"
date: "`r Sys.Date()`"
output:
    BiocStyle::html_document:
    toc: true
vignette: >
    %\VignetteIndexEntry{Example_simulated}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r}
library(gscreend)
library(SummarizedExperiment)
```

# Introduction

Pooled CRISPR perturbations screens employ a library of guide RNAs that 
is transduced into a pool of cells with the aim to induce a single genetic 
perturbation in each cell. The perturbation effect is assessed by 
measuring the abunndance of each gRNA after the screen selection 
phase and comparing it to its abundance in the plasmid library. 
The main goal of the following analysis is the detection of 
essential genes, i.e. genes whose knockout reduces the cell fitness. 
The package gscreend provides a method to rank genes based on count tables.

# gscreend workflow

In order to identify essential genes starting from raw gRNA count data,
gscreend performs the following analsysis steps:
1. Input of raw gRNA counts at T0 (seuqencing of library) 
and T1 (at the end of the screen). Normalization and 
calculation of log fold changes.
2. Split log fold changes into intervals dependent on the 
initial count at T0.
3. For every interval fit a skew-normal distibution to the 
data to model the null hypothesis (via least quantile regression).
4. Based on the null model calculate p-values for every sgRNA.
5. Rank sgRNAs according to p-value and perform robust ranking 
aggregation to calculate p-values on gene level.
6. Perform quality control of data and statistical model.

# Installation

devtools::install_github("imkeller/gscreend")

# Analysis of simulated data with gscreend

## Input data: sgRNA counts

The simulated data used in this example has been generated 
using ths simulation method available at 
https://github.com/imkeller/simulate_pooled_screen

Raw count data consists of gRNA counts in the libary sequencing 
and different replicates after the screen proliferation phase.

```{r}
raw_counts <- read.table(
    system.file("extdata", "simulated_counts.txt", package = "gscreend"),
    header=TRUE)
```

Generate a summarized experiment fromt the count data

```{r}
counts_matrix <- cbind(raw_counts$library0, 
                        raw_counts$R0_0, 
                        raw_counts$R1_0)

rowData <- data.frame(sgRNA_id = raw_counts$sgrna_id,
                    gene = raw_counts$Gene)

colData <- data.frame(samplename = c("library", "R1", "R2"),
                    # timepoint naming convention: 
                    # T0 -> reference, 
                    # T1 -> selected
                    timepoint = c("T0", "T1", "T1"))

se <- SummarizedExperiment(assays=list(counts=counts_matrix),
                    rowData=rowData, colData=colData)
```

## Run gscreend

In this step a gscreend experiment object is generated that 
will after the analysis contain all data related to gRNAs, 
genes and model parameters.

```{r}
pse <- createPoolScreenExp(se)
```

Run gscreend

```{r}
pse_an <- RunGscreend(pse, 0.1, 0.9, 0.1)

```

## Quality control

gscreend provides basic quality control functions for inspection 
of replicate correlation for example.

```{r}
plotReplicateCorrelation(pse_an)
```

Model parameters from fit.

```{r}
plotModelParameters(pse_an) 
```

# Results

Essential genes results table

```{r}
ResultsTable(pse_an, direction = "negative")
```


